\Abstract{
Navigation across multiple floors is necessary to expand the range of applications for mobile robots. In particular, for service robots in hospitals, the ability to perform a wide range of tasks across multiple floors could address labor shortages. To move in areas where other humans work, the path must be straight and predictable. The goal of this work is to find a solution for multi-floor navigation in arbitrarily complex environments while avoiding disturbance of public space. For this purpose, a hierarchical graph (H-graph) of the environment is created and the optimal path is found. At the lowest level, a deterministic roadmap with straight paths is created by iteratively finding the largest interior rectangles (ILIR) and connecting them to doors or elevators. The whole navigation over several floors is coordinated with a behavior tree (BT). Comparison with common algorithms such as A*, PRM, and RRT on established benchmarks shows that the ILIR algorithm is the fastest planner and outperforms the next best planner on the new metric measuring public space disturbance by up to 0.72 \% points. A qualitative comparison with related work shows that the generated roadmap is straighter and more predictable for humans. Finally, the integrated system was extensively tested in a simulated hospital environment with three different robots. The contributions of this work are a hierarchical planner, the ILIR planner for roadmap generation, and the integration of these approaches into the widely used Nav2 stack for ROS 2 Foxy.
}